### 2009-04-22 docs/flowchart/flowchart.tex
minor cleanup
starting to merge spherical base state stuff into flowchart

\\ \\

@@

\item[plane-parallel:] We discretize the base state density evolution 
equation (paper III; eq.~[29]) without the $\etarho$ term,
\begin{eqnarray}
\frac{\partial \rhozero}{\partial t} &=& - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{eqnarray}
to compute the new base state density,
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
         \left( {{\rho}_0}^{\outp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right ]
         \nonumber \enskip,
\end{eqnarray}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, ${{\rho}_0}^{\outp,\nph}$, are found via the
procedure described in paper II, Appendix A.

@@

\item[plane-parallel:] We discretize
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ {(\rho h)_0}^{\nph} {w_0}^{\inp}\right]_{j+\myhalf} -  
         \left[ {(\rho h)_0}^{\nph} {w_0}^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
is the process by which we adjust the base state density to account for
large-scale mixing.  It can also be used to adjust the base state enthalpy
to account for large-scale mixing, reactions, and thermal diffusion.

@@

%--------------------------------------------------------------------------
% STEP 8.1
%--------------------------------------------------------------------------
\item Diffuse the enthalpy through a time interval of $\dt$.  Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.  The update is given by
\begin{eqnarray}
(\rho h)^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
which is numerically implemented as a diffusion equation for $h^{(2)}$,
\begin{eqnarray}
\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
\item Update the temperature using the equation of state.  For plane-parallel we
update using
\begin{equation}
T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) \enskip ,
\end{equation}
but in spherical, we set {\tt use\_tfromp = T}, and we use
\begin{equation}
T^{(2)} = T\left(\rho^{(2)}, {p_0}^{(2)}, X_k^{(2)}\right) \enskip .
\end{equation}
The latter has the effect of completely decoupling the enthalpy equation from the 
evolution.

@@

\item Define
\begin{eqnarray}
 \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, {X_k}^{n+1}\right) \right] \enskip , \\ 
 {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
\end{eqnarray}

>>>>

\section{Shorthand Functions}

@@

\item[plane-parallel:] We discretize the base state density evolution 
equation (paper III; eq.~[29]) without the $\etarho$ term,
\begin{equation}
\frac{\partial \rhozero}{\partial t} = - \nablab \cdotb \left( \rhozero w_0 \er \right)
\enskip ,
\label{eq:rho0upd_new}
\end{equation}
to compute the new base state density,
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp} - \frac{\dt}{\dr} 
\left [ \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
         \left( \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right ]
         \nonumber \enskip,
\end{equation}
where $j$ refers to the one-dimensional index in the radial direction.
The interface states, $\rho_0^{\outp,\nph}$, are found via the
procedure described in paper II, Appendix A.\\

\item[spherical:] The base state density update now includes the area factors in the 
divergences:
\begin{equation}
\rho_{0,j}^{\outp} = \rho_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left [ \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j+\myhalf} -  
          \left( r^2 \rho_0^{\outp,\nph} w_0^{\inp}\right)_{j-\myhalf} \right] \enskip. 
\end{equation}

@@

\item[plane-parallel:] We discretize
\begin{equation}
\frac{\partial (\rho h)_0}{\partial t} = -\nabla\cdot\left[(\rho h)_0w_0\er\right] 
+ \psi \enskip ,
\end{equation}
to compute the new base state enthalpy,
\begin{equation}
(\rho h)_{0,j}^{\outp} = (\rho h)_{0,j}^{\inp} - \frac{\dt}{\Delta r}
\left\{ \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j+\myhalf} -  
         \left[ (\rho h)_0^{\nph} w_0^{\inp}\right]_{j-\myhalf} \right\}
+ \dt\psi_j^{\inp} \enskip ,
\end{equation}
where the interface states are again found via the procedure described
in paper II, Appendix A.

@@

\item {\bf Correct Base}$[\rho_0^{\inp}, \rho^{\inp}] \rightarrow [\rho_0^{\outp}]$\\
is the process by which we adjust the base state density to account for
large-scale mixing.  It is also used to adjust the base state enthalpy
to account for large-scale mixing, reactions, and thermal diffusion.

@@

%--------------------------------------------------------------------------
% STEP 8.1
%--------------------------------------------------------------------------
\item Diffuse the enthalpy through a time interval of $\dt$.  Compute $\kth^{(2),\star}, c_p^{(2),\star}$, and $\xi_k^{(2),\star}$, from $\rho^{(2),\star}, T^{(2),\star}$, and $X_k^{(2),\star}$ as inputs to the equation of state.  The update is given by
\begin{eqnarray}
(\rho h)^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\left(\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla h^{(2)} + \frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
which is numerically implemented as a diffusion equation for $h^{(2)}$,
\begin{eqnarray}
\left(\rho^{(2)} - \frac{\dt}{2}\nabla\cdot\frac{\kth^{(2),\star}}{c_p^{(2),\star}}\nabla\right)h^{(2)} &=& (\rho h)^{(1a)} + \frac{\dt}{2}\nabla\cdot\frac{\kth^{(1)}}{c_p^{(1)}}\nabla h^{(1)}\nonumber\\
&&- \frac{\dt}{2}\sum_k\nabla\cdot\left(\frac{\xi_k^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla X_k^{(2)} + \frac{\xi_k^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla X_k^{(1)}\right)\nonumber\\
&&- \frac{\dt}{2}\nabla\cdot\left(\frac{h_p^{(2),\star}\kth^{(2),\star}}{c_p^{(2),\star}}\nabla p_0^{(2)} + \frac{h_p^{(1)}\kth^{(1)}}{c_p^{(1)}}\nabla p_0^{(1)}\right),
\end{eqnarray}
\item Update the temperature using the equation of state.  For plane-parallel we
update using
\begin{equation}
T^{(2)} = T\left(\rho^{(2)}, h^{(2)}, X_k^{(2)}\right) \enskip ,
\end{equation}
but in spherical, we set {\tt use\_tfromp = T}, and we use
\begin{equation}
T^{(2)} = T\left(\rho^{(2)}, p_0^{(2)}, X_k^{(2)}\right) \enskip .
\end{equation}
The latter has the effect of completely decoupling the enthalpy equation from the 
evolution.

@@

\item Define
\begin{eqnarray}
 \gammabar^{n+1}    &=& {\rm{\bf Avg}}\left[\Gamma_1\left(\rho^{n+1}, p_0^{n+1}, X_k^{n+1}\right) \right] \enskip , \\ 
 {\beta   }_0^{n+1}    &=& \beta   \left(\rho_0^{n+1}, p_0^{n+1},   \gammabar^{n+1}\right) \enskip .
\end{eqnarray}

### 2009-04-22 docs/spherical_basestate/basestate.tex
shrunk the image so the caption is readable
removed a marginpar we've addressed
starting to merge spherical base state stuff into flowchart

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{i-\half} \nonumber\\
&=& 
- \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
   - \left[ \frac{4 \ow \rho_0 g}{r} \right]_{i-\half}
   - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip , \nonumber
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{i  } \frac{[ (r^2 \dw)_{i+\half} - (r^2 \dw)_{i-\half} ]}{\Delta r}
-\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \frac{[ (r^2 \dw)_{i-\half} - (r^2 \dw)_{i-\thalf}]}{\Delta r} \right\} \nonumber \\
&&  - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \dw)_{i-\half}  \nonumber \\
&=&   \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
 - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
\end{eqnarray} \MarginPar{note in the RHS, we do a difference of cell-centered $\eta$'s, but $\eta$ is defined on edges -- what should we do?}

@@

\item {\bf Density Update:}

The base state density update now includes the area factors in the divergences:
\begin{eqnarray}
\rho_{0,j}^{\outp} &=& \rho_{0,j}^{\inp}
 - \frac{1}{r_j^2} \frac{\dt}{\dr} 
 \left [ \left( {{r^2 \rho}_0}^{\inp,\nph} {w_0}^{\inp}\right)_{j+\myhalf} -  
          \left( {{r^2 \rho}_0}^{\inp,\nph} {w_0}^{\inp}\right)_{j-\myhalf} \right]   \enskip. 
\end{eqnarray}
As before, the interface states are found via the procedure described in
paper II, Appendix A.

@@

\begin{figure*}
\begin{center}
\epsffile{\sphericalfigpath/spherical_adjust_768.eps}
\end{center}
\caption{\label{fig:spherical768} Hydrostatic adjustment of a
spherically symmetric white dwarf with self-gravity.  The gray line
represents the initial model;  all other lines are after 10~s of heating.
The solid black line is the fully compressible solution, the dotted line is the
low Mach number solution with a CFL number of 0.5,  and the dashed line is
the low Mach number solution with a CFL number of 0.1.  All
simulations used 768 equally spaced zones.  We see excellent agreement
between the compressible and low Mach number models.  The only
differences appear at the top of the atmosphere, where the outer
boundary condition can influence the results.}
\end{figure*}

>>>>

\noindent We discretize this elliptic equation in the radial dimension as:
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i} -
\left[ \frac{\gammabar p_0}{r^2} \frac{\partial (r^2 \dw)}{\partial r} \right]_{i-1} \right\}
+ \left[ \frac{4 (r^2 \dw) \rho_0 g}{r^3} \right]_{i-\half} \nonumber\\
&=& 
- \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
\left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
   - \left[ \frac{4 \ow \rho_0 g}{r} \right]_{i-\half}
   - \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip , \nonumber
\end{eqnarray}
where we have chosen to solve for $(r^2 \dw)$ rather than $\dw.$
Then, using hydrostatic equilibrium, we expand this to 
\begin{eqnarray}
&& \frac{1}{\Delta r} \left\{
 \left( \frac{\gammabar p_0}{r^2}\right)_{i  } \frac{[ (r^2 \dw)_{i+\half} - (r^2 \dw)_{i-\half} ]}{\Delta r}
-\left( \frac{\gammabar p_0}{r^2}\right)_{i-1} \frac{[ (r^2 \dw)_{i-\half} - (r^2 \dw)_{i-\thalf}]}{\Delta r} \right\} \nonumber \\
&&  - \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \dw)_{i-\half}  \nonumber \\
&=&   \left\{ \frac{4}{r_{i-\half} ^3} \frac{{p_0}_i - {p_0}_{i-1}}{\dr} \right\} (r^2 \ow)_{i-\half}
 - \frac{g_{i-\half}}{r_{i-\half}^2 \Delta r} 
  \left[ \left( r^2 \etarho \right)_{i} - \left( r^2 \etarho  \right)_{i-1} \right] 
- \left[  4 \pi G \rhozero \etarho   \right]_{i-\half} \enskip.
\end{eqnarray}

@@

@@

\begin{figure*}
\begin{center}
\epsfxsize=4.0in
\epsffile{\sphericalfigpath/spherical_adjust_768.eps}
\end{center}
\caption{\label{fig:spherical768} Hydrostatic adjustment of a
spherically symmetric white dwarf with self-gravity.  The gray line
represents the initial model;  all other lines are after 10~s of heating.
The solid black line is the fully compressible solution, the dotted line is the
low Mach number solution with a CFL number of 0.5,  and the dashed line is
the low Mach number solution with a CFL number of 0.1.  All
simulations used 768 equally spaced zones.  We see excellent agreement
between the compressible and low Mach number models.  The only
differences appear at the top of the atmosphere, where the outer
boundary condition can influence the results.}
\end{figure*}
